---
output: github_document
bibliography: inst/REFERENCES.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(inspector)
```

# inspector: Validation of Arguments and Objects in User-Defined Functions

<!-- badges: start -->
[![CRAN
status](https://www.r-pkg.org/badges/version/inspector)](https://CRAN.R-project.org/package=inspector)
[![Build Status](https://travis-ci.com/pedro-teles-fonseca/inspector.svg?branch=master)](https://travis-ci.com/pedro-teles-fonseca/inspector)
[![R build status](https://github.com/pedro-teles-fonseca/inspector/workflows/R-CMD-check/badge.svg)](https://github.com/pedro-teles-fonseca/inspector/actions)
![pkgdown](https://github.com/pedro-teles-fonseca/inspector/workflows/pkgdown/badge.svg)
[![codecov](https://codecov.io/gh/pedro-teles-fonseca/inspector/branch/master/graph/badge.svg?branch=master&kill_cache=1)](https://codecov.io/gh/pedro-teles-fonseca/inspector)
[![MIT license](https://img.shields.io/badge/License-MIT-brightgreen.svg)](https://lbesson.mit-license.org/)
<!-- badges: end -->

## Overview

The `inspector` package provides utility functions that implement and automate common sets of validation tasks, namely:

  - `inspect_prob()` checks if an object is a numeric vector of valid probability values.
  
  - `inspect_log_base()` checks if an object is a valid logarithmic base.

  - `inspect_true_or_false()` checks if an object is a non-missing logical value. 

  - `inspect_bfactor()` checks if an object is a numeric vector of valid Bayes factors values.
  
  - `inspect_bfactor_log()` checks if an object is a numeric vector of valid logarithmic Bayes factors values. 
  
  - `inspect_bfactor_scale()` validates Bayes factor interpretation scales (from the [`pcal`](https://pedro-teles-fonseca.github.io/pcal/) package).

  - `inspect_categories()` validates factor levels.
  
  - `inspect_character()` validates character vectors.
  
  - `inspect_character_match()` validates character values with predefined allowed values.
  
  - `inspect_data_dichotomous()` validates dichotomous data

  - `inspect_data_categorical()` and `inspect_data_cat_as_dichotomous()` validate categorical data.

  - `inspect_par_bernoulli()` validates parameters for the Bernoulli and Binomial distributions.
  
  - `inspect_par_multinomial()` validates parameters for the Multinomial distribution.
  
  - `inspect_par_beta()` validates parameters for the Beta distribution.

  - `inspect_par_dirichlet()` validates parameters for the Dirichlet distribution.

  - `inspect_par_haldane()` validates parameters for the Haldane distribution.
  
These functions are particularly useful to validate inputs, intermediate objects and output values in user-defined functions, resulting in tidier and less verbose functions.
  
## Installation

The development version of `inspector` can be installed from
[GitHub](https://github.com/) using the [`devtools`](https://github.com/r-lib/devtools) package:

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("pedro-teles-fonseca/inspector")
```

## Usage

Imagine we want to write a function that simulates `n` flips of the same coin. Assuming that `bias` is the probability of the "heads" outcome:

```{r}

set.seed(123)

flip_coins <- function(n, bias){ 
  
  sample(x = c("heads", "tails"), size = n, replace = TRUE)
}

flip_coins(n = 5, bias = 0.5)
```

Since `bias` is a probability, it is natural that we require `flip_coins()`
to only accept values of `bias` between 0 and 1. Furthermore, we may want to ensure that `bias` is not null, not missing, and is a numeric vector of length 1. This results an a quite verbose function body:

```{r}

set.seed(123)

flip_coins <- function(n, bias){
  
  if(is.null(bias)){
    stop(paste("Invalid argument: bias is NULL."))
  }
  if(any(isFALSE(is.atomic(bias)), isFALSE(is.vector(bias)))){
    stop(paste("Invalid argument: bias must be an atomic vector."))
  }
  if(isFALSE(length(bias) == 1)){
    stop(paste("Invalid argument: bias must be of length 1."))
  }
  if(is.na(bias)){
    stop(paste("Invalid argument: bias is NA or NaN."))
  }
  if(isFALSE(is.numeric(bias))){
    stop(paste("Invalid argument: bias must be numeric."))
  }
  if(any(bias >= 1, bias <= 0)) {
    stop(paste("Invalid argument: bias must be in the (0, 1) interval."))
  }
  
  sample(x = c("heads", "tails"), size = n, replace = TRUE)
}

flip_coins(n = 5, bias = 0.5)
```

The `inspector` package was built to automate this kind of validation task. In the `flip_coins()` example, to perform an equivalent validation of inputs we can use `inspect_par_bernoulli`, since `bias` is the parameter of a Bernoulli distribution:

```{r}

set.seed(123)

flip_coins <- function(n, bias){
  
  inspect_par_bernoulli(bias)
  
  sample(x = c("heads", "tails"), size = n, replace = TRUE)
}

flip_coins(n = 5, bias = 0.5)
```

This results in a tidier function body since the validation of `bias` is abstracted away from the body of the function.

Now imagine we want to implement equation 4 from @bergerDelampady1987, a formula that calculates posterior probabilities using prior probabilities and Bayes factors as input. In this case we need to validate a vector of Bayes factors, lets call it `bf`, and a vector of prior probabilities, lets call it `prior_prob`. Since `bf` is expected to contain valid Bayes factor values, we need to ensure that only non-empty numeric vectors, containing only non-negative values, are accepted. Since `prior_prob` is a vector of probabilities, we need to check if it is a non-empty numeric vector containing only values between 0 and 1. Since we are now validating two inputs, the function body would be even more verbose than in the `flip_coins()` example:

```{r}
bfactor_to_prob <- function(bf, prior_prob = .5) {

  if(is.null(bf)){
    stop(paste("Invalid argument: bf is NULL."))
  }
  if(any(isFALSE(is.atomic(bf)), isFALSE(is.vector(bf)))){
    stop(paste("Invalid argument: bf must be an atomic vector."))
  }
  if(length(bf) == 0){
    stop(paste("Invalid argument: bf is empty."))
  }
  if(all(is.na(bf))){
    stop(paste("Invalid argument: all elements of bf are NA or NaN."))
  }
  if(isFALSE(is.numeric(bf))){
    stop(paste("Invalid argument: the type of bf must be numeric."))
  }
  if(any(bf[!is.na(bf)] < 0)){
    stop(paste("Invalid argument: all elements of bf must be non-negative."))
  }
  if(is.null(prior_prob)){
    stop(paste("Invalid argument:", output_name, "is NULL."))
  }
  if(any(isFALSE(is.atomic(prior_prob)), isFALSE(is.vector(prior_prob)))){
    stop(paste("Invalid argument:", output_name, "must be an atomic vector."))
  }
  if(length(prior_prob) == 0){
    stop(paste("Invalid argument:", output_name, "is empty."))
  }
  if(all(is.na(prior_prob))){
    stop(paste("Invalid argument: all elements of", output_name, "are NA or NaN."))
  }
  if(isFALSE(is.numeric(prior_prob))){
    stop(paste("Invalid argument: the type of", output_name, "must be numeric."))
  }
  if(any(prior_prob[!is.na(prior_prob)] < 0, prior_prob[!is.na(prior_prob)] > 1)){
    stop(paste("Invalid argument: all elements of",  output_name, "must be in the [0, 1] interval."))
  }

  (1 + (1 - prior_prob) / prior_prob * (1 / bf)) ^(-1)
}

bfactor_to_prob(c(2.1, 0.5, 11))

```

Now lets use `inspector` instead. To perform an equivalent validation of inputs we can use `inspect_bfactor()` to validate `bf` and `inspect_prob()` to validate `prior_prob`:

```{r}
bfactor_to_prob <- function(bf, prior_prob = .5) {

  inspect_bfactor(bf)
  inspect_prob(prior_prob)

  (1 + (1 - prior_prob) / prior_prob * (1 / bf)) ^(-1)
}

bfactor_to_prob(c(2.1, 0.5, 11))
```


## Getting Help

If you find a bug, please file an issue with a minimal reproducible
example on [GitHub](https://github.com/pedro-teles-fonseca/inspector).
Feature requests are also welcome. You can contact me at
<pedro.teles.fonseca@phd.iseg.ulisboa.pt>.

## References
