---
output: github_document
bibliography: inst/REFERENCES.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(inspector)
```

# inspector: Validation of Arguments and Objects in User-Defined Functions

<!-- badges: start -->
[![Build Status](https://travis-ci.com/pedro-teles-fonseca/inspector.svg?branch=master)](https://travis-ci.com/pedro-teles-fonseca/inspector)
[![R build status](https://github.com/pedro-teles-fonseca/inspector/workflows/R-CMD-check/badge.svg)](https://github.com/pedro-teles-fonseca/inspector/actions)
![pkgdown](https://github.com/pedro-teles-fonseca/inspector/workflows/pkgdown/badge.svg)
[![codecov](https://codecov.io/gh/pedro-teles-fonseca/inspector/branch/master/graph/badge.svg)](https://codecov.io/gh/pedro-teles-fonseca/inspector)
[![MIT license](https://img.shields.io/badge/License-MIT-brightgreen.svg)](https://lbesson.mit-license.org/)
<!-- badges: end -->

## Overview

The `inspector` package provides utility functions that implement and automate common sets of validation tasks, namely:

  - `inspect_prob()` checks if an object is a numeric vector of valid probability values.

  - `inspect_bfactor()` checks if an object is a numeric vector of valid Bayes factors values.
  
  - `inspect_bfactor_log()` checks if an object is a numeric vector of valid logarithmic Bayes factors values. 
  
  - `inspect_log_base()` checks if an object is a valid logarithmic base.

  - `inspect_true_or_false()` checks if an object is a non-missing logical value. 

  - `inspect_scale` check if an object is a valid Bayes factor interpretation scale.

  - `inspect_categories()` validates factor levels.
  
  - `inspect_character()` validates character vectors.
  
  - `inspect_character_match()` validates character values with predefined allowed values.
  
  - `inspect_data_dichotomous()` validates dichotomous data

  - `inspect_data_categorical()` and `inspect_data_multinom_as_bern()` validate categorical data.

  - `inspect_par_bernoulli()` validates parameters for the Bernoulli and Binomial distributions.
  
  - `inspect_par_multinomial()` validates parameters for the Multinomial distribution.
  
  - `inspect_par_beta()` validates parameters for the Beta distribution.

  - `inspect_par_dirichlet()` validates parameters for the Dirichlet distribution.

  - `inspect_par_haldane()` validates parameters for the Haldane distribution.
  
  
These functions are particularly useful to validate inputs, intermediate objects and output values in user-defined functions, resulting in tidier and less verbose functions.
  
## Installation

The development version of `inspector` can be installed from
[GitHub](https://github.com/) using the `devtools` package:

``` r
# install.packages("devtools")
devtools::install_github("pedro-teles-fonseca/inspector")
```

## Usage

```{r}

set.seed(123)

flip_coin <- function(n, bias){
  
  sample(c("heads", "tails"), size = n, replace = TRUE)

}

flip_coin(n = 5, bias = 0.5)
```

```{r}

set.seed(123)

flip_coin <- function(n, bias){
  
  if(is.null(bias)){
    stop(paste("Invalid argument: bias is NULL."))
  }
  if(any(isFALSE(is.atomic(bias)), isFALSE(is.vector(bias)))){
    stop(paste("Invalid argument: bias must be an atomic vector."))
  }
  if(isFALSE(length(bias) == 1)){
    stop(paste("Invalid argument: bias must be of length 1."))
  }
  if(is.na(bias)){
    stop(paste("Invalid argument: bias is NA or NaN."))
  }
  if(isFALSE(is.numeric(bias))){
    stop(paste("Invalid argument: bias must be numeric."))
  }
  if(any(bias >= 1, bias <= 0)) {
    stop(paste("Invalid argument: bias must be in the (0, 1) interval."))
  }
  
  sample(c("heads", "tails"), size = n, replace = TRUE)
}

flip_coin(n = 5, bias = 0.5)
```

```{r}

set.seed(123)

flip_coin <- function(n, bias){
  
  inspect_par_bernoulli(bias)
  
  sample(c("heads", "tails"), size = n, replace = TRUE)

}

flip_coin(n = 5, bias = 0.5)
```

turns Bayes factors into posterior probabilities using a formula from @bergerDelampady1987

## Getting Help

If you find a bug, please file an issue with a minimal reproducible
example on [GitHub](https://github.com/pedro-teles-fonseca/inspector).
Feature requests are also welcome. You can contact me at
<pedro.teles.fonseca@phd.iseg.ulisboa.pt>.

## References
